Blame and Coercion: Together Again for the Second Time

### 2019-10-10 paper.tex
standardize type notations and side conditions

\newtheorem{theorem}{Theorem}[]
\newtheorem{proposition}{Proposition}[]
\newtheorem{corollary}{Corollary}[]
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\redrule}[3]{#1 & \longrightarrow_B & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
\newcommand{\sOOinspect}[3]{\mathtt{Eval} \; #1 \; #2 \; #3}
\newcommand{\sOOreturn}[2]{\mathtt{Apply} \; #2 \; #1}
\newcommand{\sOOhalt}[1]{\mathtt{Halt} \; #1}
\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\iota}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{tt}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{#1 \; #2}
\newcommand{\eOOcons}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\eOOcar}[1]{\mathtt{car} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{cdr} \; #1}
\newcommand{\eOOinl}[1]{\mathtt{inl} \; #1}
\newcommand{\eOOinr}[1]{\mathtt{inr} \; #1}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{tt}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \overset{#2}{\Rightarrow} #3}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{tt}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}
\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #1 \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{tt}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}

@@

	Term typing
	\fbox{$ \Gamma \vdash e \vdash T $}
	\begin{gather*}
		\inference{
			S \sim T & \Gamma \vdash e : S 
		}{
			\judgetype{\Gamma}{\eOOcast{e}{T_1}{l}{T_2}}{T_2}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\eOOblame{l}}{T}
		}
	\end{gather*}

@@

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all blames in 
all lazy strategies -- casting a value to a shallowly inconsistent type leads 
to a blame immediately. Shallow-consistency is also reflexive, symmetric, and 
not transitive.

Let $ e $ ranges over terms, including all terms from $ \mathtt{STLC+} $, 
casts, and blames. Unlike $ \mathtt{STLC+} $, we annotate the 
co-domain of lambda abstractions explicitly. 
The end of the figure shows the extra typing rules.

@@

	Continuation typing \fbox{$ \kappa : T_1 \Longrightarrow T_2 $}
	\begin{gather*}
	\dots \quad
	\inference{
		c : T_1 \Longrightarrow T_2 &
		\kappa : T_2 \Longrightarrow T_3
	}{
		\langle c \rangle \kappa : T_1 \Longrightarrow T_3
	}
	\end{gather*}

	Reduction \fbox{$ s_1 \longrightarrow_B s_2 $}
	\[
	\begin{array}{rclr}

	& \vdots \\
	\redrule{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{\langle\cOOcast{T_1}{l}{T_2}\rangle\kappa}
	}{}
	\redrule{
		\sOOreturn{v_1}{(\mathtt{app_2} \; \vOOcast{v_2}{
				\cOOcast{\POOfun{T_{11}}{T_{12}}}{l}{\POOfun{T_{21}}{T_{22}}}
			} \; \kappa)}
	}{
		\sOOreturn{v_1}{
		\langle\cOOcast{T_{21}}{l}{T_{11}}\rangle
		(\mathtt{app_2} \; v_2 \; 
		\langle\cOOcast{T_{12}}{l}{T_{22}}\rangle \kappa)}
	}{}

@@

	Apply cast
	\[
	\begin{array}{rclr}
	\funrule{
		applyCast(\cOOcast{\star}{l}{\star},v)
	}{
		\rOOsucc{v}
	}{}
	\funrule{
		applyCast(\cOOcast{\star}{l}{P_2},\vOOcast{v}{\cOOcast{P_1}{l'}{\star}})
	}{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{}
	\funrule{
		applyCast(\cOOcast{P}{l}{\star},v)
	}{
		\rOOsucc{\vOOcast{v}{\cOOcast{P}{l}{\star}}}
	}{}
	\funrule{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{
		\rOOsucc{\vOOcast{v}{\cOOcast{P_1}{l}{P_2}}}
	}{P_1 \smile P_2}
	\funrule{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{
		\rOOfail{l}
	}{\neg P_1 \smile P_2}

@@

	\caption{Dynamic semantics of the \lazyD{} Blame Calculus}
	\label{ld-blame-cek}
\end{figure}

Fig.~\ref{ld-blame-cek} defines the dynamic semantcis of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}). 
We choose CEK machine because to achieve space efficiency, we need to change 
continuations. Thus, abstract machines that include continuation in states is 
quite convenient for us. And CEK is one of the simplest machine of this kind.

Let $ c $ ranges over casts. A cast is a triple of a type, a label, and a type.

Let $ v $ ranges over values. A value is either a function, the 
element of unit, a pair, a left injection, a right injection, or a casted value.
The cast around a casted value is not arbitrary, as shown by value typing: the 
source type must be a pre-type, and the source and the target must be 
shallow-consistent.

Let $ r $ ranges over cast results. A cast result is either a success, which 
brings a value, or a failure, which brings a blame label.

Let $ s $ ranges over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.

Let $ \kappa $ ranges over continuations. $ \mathtt{mt} $ is the top 
continuation. The following continuations correspond to expressions. For 
example, $ (\mathtt{cons_1} \; e \; \rho \; \kappa) $ is the continuation where 
we are waiting for the value of the first argument to a $ \mathtt{cons} $. And 
the last continuation, $ \langle c \rangle \kappa $ is to cast the value before 
returning to $ \kappa $.

@@

@@

	Coercion typing \fbox{$ c : T \Longrightarrow T $}
	\begin{gather*}
	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	\quad
	\inference{
		\typingHC{h}{T_1}{P_1} &
		\typingHC{m}{P_1}{P_2} &
		\typingHC{t}{P_2}{T_2}
	}{
		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	}
	\end{gather*}

	Head typing \fbox{$ \typingHC{h}{T}{P} $}
	\begin{gather*}
	\inference{}{\typingHC{\epsilon}{P}{P}}
	\quad
	\inference{}{\typingHC{?^l}{\TOOdyn}{P}}
	\end{gather*}

	Middle typing \fbox{$ \typingHC{m}{P}{P} $}
	\begin{gather*}
	\inference{}{\typingHC{\POOunit}{\POOunit}{\POOunit}}
	\\
	\inference{
		\typingHC{c_1}{T_3}{T_1} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOfun{c_1}{c_2}}{\POOfun{T_1}{T_2}}{\POOfun{T_3}{T_4}}
	}
	\\
	\inference{
		\typingHC{c_1}{T_1}{T_3} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOprod{c_1}{c_2}}{\POOprod{T_1}{T_2}}{\POOprod{T_3}{T_4}}
	}
	\\
	\inference{
		\typingHC{c_1}{T_1}{T_3} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOsum{c_1}{c_2}}{\POOsum{T_1}{T_2}}{\POOsum{T_3}{T_4}}
	}
	\end{gather*}

	Tail typing \fbox{$ \typingHC{t}{P}{T} $}
	\begin{gather*}
	\inference{}{\typingHC{\epsilon}{P}{P}} \quad
	\inference{}{\typingHC{!}{P}{\TOOdyn}} \quad
	\inference{}{\typingHC{\bot^l}{P}{T}} \quad
	\end{gather*}

@@

\figref{fig:hyper-coercion} defines the syntax of hyper-coercions. All 
definitions and results in this section are new.

Terms, types, pre-types, blame labels, and observations are as before. Let $ c 
$ ranges over hyper-coercions. A hyper-coercion is either an identity 
cast between the dynamic types, or a complex including a head, a middle, and a 
tail. Let $ h $ ranges over heads. A head is either a no-op, or a projection 
decorated with a label, to which blame is allocated if the projection fails. 
Let $ m $ ranges over middles. There is a one-to-one 
correspondence between middles and type constructors. Let $ t $ ranges over 
tails. A tail is either a no-op, an injection, or a failure. Let $ \ell $ 
ranges over no-op and labels. It is useful when we introduce composition of 
hyper-coercions. Typing rules for hyper-coercions, heads, middles, and tails 
are straightforward.

@@

	\comprule{
		\hyperCoercionI
	}{
		\hyperCoercionC{!}{b}{t}
	}{
		\hyperCoercionC{?^{l}}{b}{t}
	}{\ell = l}

@@

	\comprule{
		\hyperCoercionC{h}{b_1}{t_1}
	}{
		\hyperCoercionI
	}{
		\hyperCoercionC{h}{b_1}{!}
	}{
		t_1 \neq \bot^{\centerdot}
	}

	\comprule{
		\hyperCoercionC{h}{b_1}{t_1}
	}{
		\hyperCoercionC{\epsilon}{b_2}{t_2}
	}{
		\hyperCoercionC{h}{b'}{t'}
	}{
		t_1 \neq \bot^{\centerdot} \; \text{and} \;
		b_1 \fatsemi^{\ell} (b_2, t) = (b', t')
	}

	\comprule{
		\hyperCoercionC{h}{b_1}{t_1}
	}{
		\hyperCoercionC{?^{l'}}{b_2}{t_2}
	}{
		\hyperCoercionC{h}{b'}{t'}
	}{
		t_1 \neq \bot^{\centerdot} \; \text{and} \;
		b_1 \fatsemi^{l'} (b_2, t) = (b', t')
	}
	\end{array}
	\]

	\fbox{$ m \fatsemi^\ell (m,t) = (m,t) $}
	\[ 
	\begin{array}{rclclr}
	\comprule{U}{(U,t)}{
		(U,t)
	}{}
	\comprule{c_1 \rightarrow c_2}{(c_3 \rightarrow c_4,t)}{
		(c_3 \fatsemi^{\ell} c_1 \rightarrow c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{c_1 \times c_2}{(c_3 \times c_4,t)}{
		(c_1 \fatsemi^{\ell} c_3 \times c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{c_1 \plus c_2}{(c_3 \plus c_4,t)}{
		(c_1 \fatsemi^{\ell} c_3 \plus c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{b_1}{(b_2,t)}{
		(b_1,\bot^l)
	}{
		\ell = l \; \text{and} \;
		\neg b_1 \smile b_2 
	}
	\end{array}
	\]

	\fbox{$ seq(c,c) = c $}
	\[
	\begin{array}{rclr}
	\funrule{seq(c_1,c_2)}{
		c_1 \fatsemi^\epsilon c_2
	}{
		c_1 : T_1 \Longrightarrow T_2
		\; \text{and} \;
		c_2 : T_2 \Longrightarrow T_3
	}
	\end{array}
	\]

	\fbox{$ id( P ) = m $}
	\[
	\begin{array}{rclr}
	\funrule{id(\POOunit)}{\POOunit}{}
	\funrule{id(\POOfun{T_1}{T_2})}{
		\POOfun{id(T_1)}{id(T_2)}
	}{}
	\funrule{id(\POOprod{T_1}{T_2})}{
		\POOprod{id(T_1)}{id(T_2)}
	}{}
	\funrule{id(\POOsum{T_1}{T_2})}{
		\POOsum{id(T_1)}{id(T_2)}
	}{}
	\end{array}
	\]

@@

Functions that construct \lazyD{} Hyper-coercion are shown in \figref{fig:HC-D}.
When a language implementation want to construct a coercion, it should use 
these functions: $ cast(T_1,l,T_2) $, $ id(T) $, and $ seq(c_1,c_2) $. $ 
cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type. $ id(T) $ constructs an identity cast of $ T $. $ seq(c_1,c_2) $ 
composes two coercions where the target type of $ c_1 $ is equal to the source 
type of $ T_2 $.

\begin{figure}
	\fbox{$ applyCast(c,v) = r $}
	\[
	\begin{array}{rclr}
	\funrule{applyCast(\hyperCoercionI,\;v)}{\rOOsucc{v}}{}
	\funrule{applyCast(\hyperCoercionC{?^l}{m}{t},\;\hcvOOinj{P}{v})}{
		applyTail(t',applyMiddle(m',v))
	}{id(P)\fatsemi^l(m,t) = (m',t')}
	\funrule{applyCast(\hyperCoercionC{\epsilon}{m}{t},\;v)}{
		applyTail(t,applyMiddle(m,v))}
	{}
	\end{array}
	\]

	\fbox{$ applyMiddle(m,v) = v $}
	\[
	\begin{array}{rclr}
	\funrule{applyMiddle(\POOunit,\hcvOOtt)}{\hcvOOtt}{}
	\funrule{applyMiddle(\POOfun{c_3}{c_4},\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{
		\hcvOOfun{seq(c_3,c_1)}{\rho}{x}{e}{seq(c_2,c_4)}
	}{}
	\funrule{applyMiddle(\POOprod{c_3}{c_4},\hcvOOcons{v_1}{c_1}{v_2}{c_2})}{
		\hcvOOcons{v_1}{seq(c_1,c_3)}{v_2}{seq(c_2,c_4)}
	}{}
	\funrule{applyMiddle(\POOsum{c_3}{c_4},\hcvOOinl{v}{c_1})}{
		\hcvOOinl{v}{seq(c_1,c_3)}
	}{}
	\funrule{applyMiddle(\POOsum{c_3}{c_4},\hcvOOinr{v}{c_2})}{
		\hcvOOinr{v}{seq(c_2,c_4)}
	}{}
	\end{array}
	\]

>>>>

\newtheorem{theorem}{Theorem}[]
\newtheorem{proposition}{Proposition}[]
\newtheorem{corollary}{Corollary}[]
\newtheorem{definition}{Definition}

\newcommand{\GTLC}{\texttt{GTLC+}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\redrule}[3]{#1 & \longrightarrow_B & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
\newcommand{\sOOinspect}[3]{\mathtt{Eval} \; #1 \; #2 \; #3}
\newcommand{\sOOreturn}[2]{\mathtt{Apply} \; #2 \; #1}
\newcommand{\sOOhalt}[1]{\mathtt{Halt} \; #1}
\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\mathtt{Unit}}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{unit}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{#1 \; #2}
\newcommand{\eOOcons}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\eOOcar}[1]{\mathtt{fst} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{snd} \; #1}
\newcommand{\eOOinl}[1]{\mathtt{inl} \; #1}
\newcommand{\eOOinr}[1]{\mathtt{inr} \; #1}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{unit}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \overset{#2}{\Rightarrow} #3}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{unit}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}
\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hckOOmt}[0]{\mathtt{mt}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}

@@

	Term typing
	\fbox{$ \judgetype{\Gamma}{e}{T} $}
	\begin{gather*}
		\inference{
			S \sim T & \Gamma \vdash e : S 
		}{
			\judgetype{\Gamma}{\eOOcast{e}{T_1}{l}{T_2}}{T_2}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\eOOblame{l}}{T}
		}
	\end{gather*}

@@

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all blames in 
all lazy strategies -- casting a value to a shallowly inconsistent type leads 
to a blame immediately. Shallow-consistency is reflexive, symmetric, but 
not transitive.

Let $ e $ ranges over terms, including all terms from $ \mathtt{STLC+} $ and, 
in addition, casts and blames. Unlike $ \mathtt{STLC+} $, we annotate the 
co-domain of lambda abstractions explicitly. 
The end of the figure shows the typing rules for the additional terms.

@@

%	Continuation typing \fbox{$ \kappa : T_1 \Longrightarrow T_2 $}
%	\begin{gather*}
%	\dots \quad
%	\inference{
%		c : T_1 \Longrightarrow T_2 &
%		\kappa : T_2 \Longrightarrow T_3
%	}{
%		\langle c \rangle \kappa : T_1 \Longrightarrow T_3
%	}
%	\end{gather*}

	Reduction \fbox{$ s_1 \longrightarrow_{D,B} s_2 $}
	\[
	\begin{array}{rclr}

	& \dots \\
	\redrule{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{\langle\cOOcast{T_1}{l}{T_2}\rangle\kappa}
	}{}
	\redrule{
		\sOOreturn{v_1}{(\mathtt{app_2} \; \vOOcast{v_2}{
				\cOOcast{\POOfun{T_{11}}{T_{12}}}{l}{\POOfun{T_{21}}{T_{22}}}
			} \; \kappa)}
	}{
		\sOOreturn{v_1}{
		\langle\cOOcast{T_{21}}{l}{T_{11}}\rangle
		(\mathtt{app_2} \; v_2 \; 
		\langle\cOOcast{T_{12}}{l}{T_{22}}\rangle \kappa)}
	}{}

@@

	Apply cast
	\[
	\begin{array}{rclr}
	\funrule{
		applyCast(\cOOcast{\star}{l}{\star},v)
	}{
		\rOOsucc{v}
	}{}
	\funrule{
		applyCast(\cOOcast{\star}{l}{P_2},\vOOcast{v}{\cOOcast{P_1}{l'}{\star}})
	}{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{}
	\funrule{
		applyCast(\cOOcast{P}{l}{\star},v)
	}{
		\rOOsucc{\vOOcast{v}{\cOOcast{P}{l}{\star}}}
	}{}
	\funrule{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{
		\rOOsucc{\vOOcast{v}{\cOOcast{P_1}{l}{P_2}}}
	}{\sidecond{P_1 \smile P_2}}
	\funrule{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{
		\rOOfail{l}
	}{\sidecond{\neg P_1 \smile P_2}}

@@

	Transitive closure of reduction \fbox{$ s \longrightarrow_{D,B}^{*} s $}\\

	Evaluation \fbox{$ e \Downarrow_{D,B} o $}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{B}^{*} 
		\sOOhalt{o}
	}{
		e \Downarrow_{D,B} o
	}
	\]

	\caption{Dynamic semantics of the \lazyD{} Blame Calculus}
	\label{machine-cekc}
\end{figure}

Fig.~\ref{machine-cekc} defines the dynamic semantcis of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}). 
We choose CEK machine because to achieve space efficiency, we need to change 
continuations. Thus, abstract machines that include continuation in states is 
quite convenient for us. And CEK is one of the simplest machine of this kind.

Let $ c $ ranges over casts. A cast is a triple of a type, a label, and a type.
Let $ v $ ranges over values. A value is either a function, the 
element of unit, a pair, a left injection, a right injection, or a casted value.
The cast around a casted value is not arbitrary, as shown by value typing: the 
source type must be a pre-type, and the source and the target must be 
shallow-consistent.
Let $ r $ ranges over cast results. A cast result is either a success, which 
brings a value, or a failure, which brings a blame label.
Let $ s $ ranges over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.
Let $ \kappa $ ranges over continuations. $ \mathtt{mt} $ is the top 
continuation. The remaining continuations correspond to expressions. For 
example, $ (\mathtt{cons_1} \; e \; \rho \; \kappa) $ is the continuation where 
we are waiting for the value of the first argument to a $ \mathtt{cons} $. And 
the last continuation, $ \langle c \rangle \kappa $ is to cast the value before 
returning to $ \kappa $.

@@

Transitive closure of reduction and evaluation are standard.

@@

	Shallow-consistency between middles
	\fbox{$ m \smile m $}

%	Coercion typing \fbox{$ c : T \Longrightarrow T $}
%	\begin{gather*}
%	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
%	\quad
%	\inference{
%		\typingHC{h}{T_1}{P_1} &
%		\typingHC{m}{P_1}{P_2} &
%		\typingHC{t}{P_2}{T_2}
%	}{
%		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
%	}
%	\end{gather*}
%	
%	Head typing \fbox{$ \typingHC{h}{T}{P} $}
%	\begin{gather*}
%	\inference{}{\typingHC{\epsilon}{P}{P}}
%	\quad
%	\inference{}{\typingHC{?^l}{\TOOdyn}{P}}
%	\end{gather*}
%	
%	Middle typing \fbox{$ \typingHC{m}{P}{P} $}
%	\begin{gather*}
%	\inference{}{\typingHC{\POOunit}{\POOunit}{\POOunit}}
%	\\
%	\inference{
%		\typingHC{c_1}{T_3}{T_1} &
%		\typingHC{c_2}{T_2}{T_4}
%	}{
%		\typingHC{\POOfun{c_1}{c_2}}{\POOfun{T_1}{T_2}}{\POOfun{T_3}{T_4}}
%	}
%	\\
%	\inference{
%		\typingHC{c_1}{T_1}{T_3} &
%		\typingHC{c_2}{T_2}{T_4}
%	}{
%		\typingHC{\POOprod{c_1}{c_2}}{\POOprod{T_1}{T_2}}{\POOprod{T_3}{T_4}}
%	}
%	\\
%	\inference{
%		\typingHC{c_1}{T_1}{T_3} &
%		\typingHC{c_2}{T_2}{T_4}
%	}{
%		\typingHC{\POOsum{c_1}{c_2}}{\POOsum{T_1}{T_2}}{\POOsum{T_3}{T_4}}
%	}
%	\end{gather*}
%	
%	Tail typing \fbox{$ \typingHC{t}{P}{T} $}
%	\begin{gather*}
%	\inference{}{\typingHC{\epsilon}{P}{P}} \quad
%	\inference{}{\typingHC{!}{P}{\TOOdyn}} \quad
%	\inference{}{\typingHC{\bot^l}{P}{T}} \quad
%	\end{gather*}

@@

All definitions and results in this section are new.
\figref{fig:hyper-coercion} defines the syntax of hyper-coercions.
Terms, types, pre-types, blame labels, and observations are as before.

Let $ c $ ranges over hyper-coercions. We re-use the meta-variable $ c $ to 
stress that hyper-coercion is a cast representation. A hyper-coercion is either 
an identity cast between the dynamic types, or a complex including a head, a 
middle, and a tail. 
Let $ h $ ranges over heads. A head is either a no-op, or a projection 
decorated with a label, to which blame is allocated if the projection fails. 
Let $ m $ ranges over middles. There is a one-to-one 
correspondence between middles and type constructors. 
We generalize shallow-consistency to include middles in the natural way.
Let $ t $ ranges over tails. A tail is either a no-op, an injection, or a 
failure. 
Let $ \ell $ ranges over no-op and labels. It is useful when we introduce 
composition of hyper-coercions.

@@

	\comprule{
		\hyperCoercionI
	}{
		\hyperCoercionC{\epsilon}{b}{t}
	}{
		\hyperCoercionC{?^{l}}{b}{t}
	}{\sidecond{\ell = l}}

@@

	\comprule{
		\hyperCoercionC{h}{b_1}{t_1}
	}{
		\hyperCoercionI
	}{
		\hyperCoercionC{h}{b_1}{!}
	}{
		\sidecond{\forall l. t_1 \neq \bot^{l}}
	}

	\comprule{
		\hyperCoercionC{h}{b_1}{t_1}
	}{
		\hyperCoercionC{\epsilon}{b_2}{t_2}
	}{
		\hyperCoercionC{h}{b'}{t'}
	}{
		\sidecond{
			t_1 \neq \bot^{\centerdot} \; \text{and} \;
			b_1 \fatsemi^{\ell} (b_2, t) = (b', t')
		}
	}

	\comprule{
		\hyperCoercionC{h}{b_1}{t_1}
	}{
		\hyperCoercionC{?^{l'}}{b_2}{t_2}
	}{
		\hyperCoercionC{h}{b'}{t'}
	}{
		\sidecond{
		t_1 \neq \bot^{\centerdot} \; \text{and} \;
		b_1 \fatsemi^{l'} (b_2, t) = (b', t')
		}
	}
	\end{array}
	\]

	\fbox{$ m \fatsemi^\ell (m,t) = (m,t) $}
	\[ 
	\begin{array}{rclclr}
	\comprule{\POOunit}{(\POOunit,t)}{
		(\POOunit,t)
	}{}
	\comprule{\POOfun{c_1}{c_2}}{(\POOfun{c_3}{c_4},t)}{
		(\POOfun{c_3 \fatsemi^{\ell} c_1}{c_2 \fatsemi^\ell c_4},t)
	}{}
	\comprule{\POOprod{c_1}{c_2}}{(\POOprod{c_3}{c_4},t)}{
		(\POOprod{c_1 \fatsemi^\ell c_3}{c_2 \fatsemi^\ell c_4},t)
	}{}
	\comprule{\POOsum{c_1}{c_2}}{(\POOsum{c_3}{c_4},t)}{
		(\POOsum{c_1 \fatsemi^\ell c_3}{c_2 \fatsemi^\ell c_4},t)
	}{}
	\comprule{m_1}{(m_2,t)}{
		(m_1,\bot^l)
	}{
		\sidecond{
		\ell = l \; \text{and} \;
		\neg m_1 \smile m_2 
		}
	}
	\end{array}
	\]

	\fbox{$ seq(c,c) = c $}
	\[
	\begin{array}{rclr}
	\funrule{seq(c_1,c_2)}{
		c_1 \fatsemi^\epsilon c_2
	}{}
	\end{array}
	\]

	\fbox{$ id( P ) = m $}
	\[
	\begin{array}{rclr}
	\funrule{id(\POOunit)}{\POOunit}{}
	\funrule{id(\POOfun{T_1}{T_2})}{
		\POOfun{id(T_1)}{id(T_2)}
	}{}
	\funrule{id(\POOprod{T_1}{T_2})}{
		\POOprod{id(T_1)}{id(T_2)}
	}{}
	\funrule{id(\POOsum{T_1}{T_2})}{
		\POOsum{id(T_1)}{id(T_2)}
	}{}s
	\end{array}
	\]

@@

Functions that construct \lazyD{} hyper-coercions are shown in 
\figref{fig:HC-D}.

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hyper-coercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.

$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.

$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.

$ id(T) $ constructs an identity coercion of $ T $.

$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.
